---
title: "In-class Exercise 4"
author: "Fong Bao Xian"
date: "6 May 2023"
---

# Getting Started

## Import and Launch R packages

-   [rstatix](https://rpkgs.datanovia.com/rstatix/) - Provides a simple framework, coherent with the 'tidyverse' design philosophy, for performing basic statistical tests, including t-test, Wilcoxon test, ANOVA, Kruskal-Wallis and correlation analyses. The output of each test is automatically transformed into a tidy data frame to facilitate visualization (tibble format).

```{r}
pacman::p_load(rstatix, gt, patchwork, tidyverse, webshot2, png)
```

## Import data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# Visualising Normal Distribution

We will use the Q-Q plot, short for "quantile-quantile" plot, to examine whether the data is normally distributed. If the data is normally distributed, the points in a Q-Q plot will lie on a straight diagonal line. Conversely, if the points deviate significantly from the straight diagonal line, then it's less likely that the data is normally distributed.

```{r}
ggplot(data = exam_data,
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()
```

::: callout-note
We can see that the points deviate significantly from the straight diagonal line. This is a clear indication that the set of data is not normally distributed.
:::

# Combining Statistical Graph and Analysis Table

```{r}
qq <- ggplot(data = exam_data,
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()

sw_t <- exam_data %>% 
  shapiro_test(ENGLISH) %>% 
  gt()

tmp <- tempfile(fileext = ".png")
gtsave(sw_t, tmp)
table_png <- png::readPNG(tmp, native = TRUE)

qq + table_png
```
