---
title: "Take-home Exercise 3"
subtitle: "Detect Anomalies in Fishing-Related Companies"
author: "Fong Bao Xian"
date: "5 June 2023"
date-modified: "`r Sys.Date()`"
---

# Overview

This exercise references [Mini-Challenge 3](https://vast-challenge.github.io/2023/MC3.html) of VAST Challenge 2023. The objective of the exercise is to help FishEye International identify companies possibly engaged in illegal, unreported and unregulated (IUU) fishing through analysis of anomalous company structures.

We will focus on task 1 in the mini-challenge, which is to use visual analytics to identify anomalies in the business groups present in the knowledge graph.

# Getting Started

## Import R Packages

For this exercise, we will be using the following packages:

-   [**tidyverse**](https://www.tidyverse.org/#:~:text=The%20tidyverse%20is%20an%20opinionated,%2C%20grammar%2C%20and%20data%20structures.&text=See%20how%20the%20tidyverse%20makes,%E2%80%9CR%20for%20Data%20Science%E2%80%9D.) - a collection of packages for data science and analysis, including packages for data import, wrangling and visualisation
-   [**jsonlite**](https://cran.r-project.org/web/packages/jsonlite/index.html) - for parsing json files from/to R objects
-   [**tidygraph**](https://tidygraph.data-imaginist.com/)- for graph and network manipulation
-   [**ggraph**](https://ggraph.data-imaginist.com/) - an extension of ggplot2 for supporting relational data structures such as networks, graphs and trees
-   [**visNetwork**](https://datastorm-open.github.io/visNetwork/)- for interactive network visualisation
-   [**igraph**](https://igraph.org/) - for creating and manipulating graphs and analysing networks
-   [**RColorBrewer**](https://r-graph-gallery.com/38-rcolorbrewers-palettes.html) - offers colour palettes for R
-   [**tidytext**](https://cran.r-project.org/web/packages/tidytext/vignettes/tidytext.html) - for text mining and analysis
-   [**DT**](https://rstudio.github.io/DT/#:~:text=The%20R%20package%20DT%20provides,other%20features%20in%20the%20tables.) - for displaying tables in R with filtering, pagination and sorting features

We will use `pacman::p_load` to install (if the packages are not yet installed) and load the packages into the R environment in a single command.

```{r}
pacman::p_load(tidyverse, jsonlite, tidygraph, ggraph, visNetwork, igraph, RColorBrewer, tidytext, DT)
```

## Import Data

We will use `fromJSON()` of **jsonlite** package to import *MC3.json* into the R environment.

```{r}
mc3 <- fromJSON("data/MC3/MC3.json")
```

# Data Preparation

## Extracting all edges

The following code chunk is used to extract *edges* data table from *mc3* list object and save the output in a tibble data frame object called *mc3_edges*. It is noted that some sources in the edges are a list of entities, hence we will use the function `unnest()` to change the list to multiple rows.

```{r}
mc3_edges <- as_tibble(mc3$links) %>% 
  unnest(source) %>% 
  distinct() %>% 
  mutate(source = as.character(source),
         target = as.character(target),
         type = as.character(type)) %>% 
  group_by(source, target, type) %>% 
  summarise(weights = n()) %>% 
  filter(source != target) %>%
  ungroup()

head(mc3_edges)
```

::: callout-note
## Things to learn from code

-   `unnest()` is used to expand a list-column of data frames into rows
-   `distinct()` is used to ensure that there will be no duplicated records.
-   `mutate()` and `as.character()` are used to convert the field data type from list to character.
-   `group_by()` and `summarise()` are used to count the number of unique links.
-   the `filter(source!=target)` is to ensure that no record with similar source and target.
:::

We will further examine the edges below, particularly the *type* field, and we can see that there are two types of edges.

```{r}
ggplot(data = mc3_edges, aes(x = type)) +
  geom_bar()
```

## Extracting all nodes

The following code chunk is used to extract *nodes* data table from *mc3* list object and save the output in a tibble data frame object called *mc3_nodes*. Note that the format of *nodes* inside *mc3* is in list format and not in dataframe format, thus it is not sufficient that we convert the *nodes* to tibble format. We still need to perform additional wrangling below.

It is also noted that there are some nodes featured in the edges but not in the nodes from the *mc3* list object. For consistency, we will combine both sets of nodes to give the complete nodes df.

```{r}
# extract all nodes from graph
mc3_nodes <- as_tibble(mc3$nodes) %>% 
  mutate(country = as.character(country),
         id = as.character(id),
         product_services = as.character(product_services),
         revenue_omu = as.numeric(as.character(revenue_omu)),
         type = as.character(type)) %>%
  select(id, country, type, revenue_omu, product_services)

# extract all nodes from edges
id1 <- mc3_edges %>%
  select(source, type) %>%
  rename(id = source) %>% 
  mutate(country = NA, revenue_omu = NA, product_services = NA) %>% 
  select(id, country, type, revenue_omu, product_services)

id2 <- mc3_edges %>%
  select(target, type) %>%
  rename(id = target) %>% 
  mutate(country = NA, revenue_omu = NA, product_services = NA) %>% 
  select(id, country, type, revenue_omu, product_services)

additional_nodes <- rbind(id1, id2) %>% 
  distinct %>% 
  filter(!id %in% mc3_nodes[["id"]])

# combine all nodes
mc3_nodes_updated <- rbind(mc3_nodes, additional_nodes) %>%
  distinct()

head(mc3_nodes_updated)
```

::: callout-note
## Things to learn from code

-   `mutate()` and `as.character()` are used to convert the field data type from list to character.
-   To convert *revenue_omu* from list data type to numeric data type, we need to convert the values into character first by using `as.character()`. Then, `as.numeric()` will be used to convert them into numeric data type.
-   For `rbind()`, we will require the same columns in the datasets, hence we use `mutate()` to add on NA columns.
-   `select()` is used to re-organise the order of the fields.
:::

We will further examine the nodes below, particularly the *type* field, and we can see that there are three types of nodes.

```{r}
ggplot(data = mc3_nodes_updated, aes(x = type)) +
  geom_bar()
```

## Cleaning up duplicated nodes

We note that there are over 3,000 rows of duplicated nodes with the same id.

```{r}
mc3_nodes_updated[duplicated(mc3_nodes_updated$id),] %>% 
  arrange(id)
```

The following code chunk is used to remove these duplicated nodes if the country and type are exactly the same.

```{r}
mc3_nodes_updated <- mc3_nodes_updated %>%
  arrange(id, country, type, desc(revenue_omu)) %>%
  distinct(id, country, type, .keep_all = TRUE)
```

After removal of duplicated nodes, each id can still appear multiple times if they are tagged to multiple countries or if they are featured as multiple types of entities. We will hence aggregate all these rows together so that each id only appears once.

```{r}
mc3_nodes_master <- mc3_nodes_updated %>% 
  group_by(id) %>% 
  arrange(id, type, country) %>% 
  summarise(countries = paste0(unique(country), collapse = ", "),
            num_countries = n_distinct(country),
            types = paste0(unique(type), collapse = ", "),
            num_types = n_distinct(type),
            revenue = sum(revenue_omu))
```

## Cleaning up product_services in nodes

On close observation of *product_services*, we note that only nodes of *type* Company have meaningful *product_services* whereas for Beneficial Owner and Company Contacts, most of the *product_services* are missing. For this reason, we will only use the product_services column for nodes that are companies.

```{r}
# extract type Company
company_nodes <- mc3_nodes_updated %>% 
  filter(type == "Company")

# extract type Beneficial Owner
bo_nodes <- mc3_nodes_updated %>% 
  filter(type == "Beneficial Owner")

# extract type Company Contacts
cc_nodes <- mc3_nodes_updated %>% 
  filter(type == "Company Contacts")

# bo nodes that have meaningful product_services
bo_nodes_w_pdtsvcs <- bo_nodes %>% 
  filter(product_services != 'character(0)' & !is.na(product_services))

# cc nodes that have meaningful product_services
cc_nodes_w_pdtsvcs <- cc_nodes %>% 
  filter(product_services != 'character(0)' & !is.na(product_services))

print(paste0("BO nodes with product_services not null: ", length(bo_nodes_w_pdtsvcs$id)))
print(paste0("CC nodes with product_services not null: ", length(cc_nodes_w_pdtsvcs$id)))
```

### Simplifying product_services

As our priority are the fishing-related companies, we will categorise fishing-related companies, others and unknown. Our categorisation will be based on related keywords in the product_services column to its own category "Seafood Products".

```{r}
# simplify product_services
company_nodes_cleaned <- company_nodes %>%
  group_by(id) %>% 
  summarise(product_services = paste0(unique(product_services), collapse = ", ")) %>% 
  mutate(category = case_when(
    str_detect(tolower(product_services), "fish|seafood|salmon|tuna|cod|fillet|fish|shrimp|squid") ~ "Seafood Products",
    is.na(product_services) ~ "Unknown",
    tolower(product_services) == "unknown" ~ "Unknown",
    TRUE ~ "Others"))

# combine cleaned product services back to mc3_master_nodes
mc3_nodes_master_updated <- mc3_nodes_master %>%
  left_join(company_nodes_cleaned, by = "id")
```

# Data Visualisation

In the following sections, we will attempt to identify anomolies in the business groups and visualise their network graph to understand further details.

## Analysis of Node with Highest Betweenness Centrality

First, we will identity the node with highest betweenness centrality in the entire network graph. A node with high betweenness centrality acts as a bridge between different parts of the graph. In our context, it may be indicative of an entity that is beneficial owner of many different companies or an entity that is used by ultimate beneficial owners to link to different groups of companies.

The node with highest betweenness centrality in this case is identified to be **Wave Warriors S.A. de C.V. Express**. We will extract the first-degree and second-degree links to **Wave Warriors S.A. de C.V. Express** and visualise the interactive network chart below.

::: callout-tip
## Interactive Chart

-   The darker orange node is our main node of interest, i.e., **Wave Warriors S.A. de C.V. Express**
-   Select the id from drop-down or click on the interested entity on the network chart to view its first-degree linked nodes
-   Select the group from drop-down to view specific types of nodes. The group refers to the types of roles that the node may play in the entire network chart, e.g., Beneficial Owner or others
-   All nodes which play a role of Beneficial Owner in at least one of its relationships will be marked with a triangle shape. This makes it easier to identify multiple layers of beneficial owners where present.
:::

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 14
#| fig-height: 14

# form graph
mc3_graph <- tbl_graph(nodes = mc3_nodes_master_updated,
                       edges = mc3_edges,
                       directed = FALSE) %>% 
  mutate(betweenness_centrality = centrality_betweenness())

# extract node with highest betweenness centrality
top1_betw <- mc3_graph %>% 
  activate(nodes) %>% 
  as_tibble() %>% 
  top_n(1, betweenness_centrality) %>% 
    select(id, countries, types)

# extract lvl 1 edges
top1_betw_edges_lvl1 <- mc3_edges %>% 
  filter(source %in% top1_betw[["id"]] | target %in% top1_betw[["id"]])

# extract nodes from lvl 1 edges
id1 <- top1_betw_edges_lvl1 %>%
  select(source) %>%
  rename(id = source) %>% 
  left_join(mc3_nodes_master_updated, by = "id") %>% 
  select(id, countries, types)

id2 <- top1_betw_edges_lvl1 %>%
  select(target) %>%
  rename(id = target) %>% 
  left_join(mc3_nodes_master_updated, by = "id") %>% 
  select(id, countries, types)

additional_nodes_lvl1 <- rbind(id1, id2) %>% 
  distinct %>% 
  filter(!id %in% top1_betw[["id"]])

# extract lvl 2 edges
top1_betw_edges_lvl2 <- mc3_edges %>% 
  filter(source %in% additional_nodes_lvl1[["id"]] | target %in% additional_nodes_lvl1[["id"]])

# extract nodes from lvl 1 edges
id1 <- top1_betw_edges_lvl2 %>%
  select(source) %>%
  rename(id = source) %>% 
  left_join(mc3_nodes_master_updated, by = "id") %>% 
  select(id, countries, types)

id2 <- top1_betw_edges_lvl2 %>%
  select(target) %>%
  rename(id = target) %>% 
  left_join(mc3_nodes_master_updated, by = "id") %>% 
  select(id, countries, types)

additional_nodes_lvl2 <- rbind(id1, id2) %>% 
  distinct %>% 
  filter(!id %in% top1_betw[["id"]] & !id %in% additional_nodes_lvl1[["id"]])

# combine all nodes
top1_betw_nodes <- rbind(top1_betw, additional_nodes_lvl1, additional_nodes_lvl2) %>%
  distinct()

# combine all edges
top1_betw_edges <- rbind(top1_betw_edges_lvl1, top1_betw_edges_lvl2) %>% 
  distinct()

# colur palatte for betweenness centrality colours
sw_colors <- colorRampPalette(brewer.pal(3, "RdBu"))(3)

# customise edges for plotting
top1_betw_edges <- top1_betw_edges %>% 
  rename(from = source,
         to = target) %>% 
  mutate(title = paste0("Type: ", type), # tooltip when hover over
         color = "#0085AF") # color of edge

# customise nodes for plotting
top1_betw_nodes <- top1_betw_nodes %>% 
  rename(group = types) %>% 
  mutate(id.type = ifelse(id == top1_betw[["id"]], sw_colors[1], sw_colors[2])) %>%
  mutate(title = paste0(id, "<br>Group: ", group), # tooltip when hover over
         size = 30, # set size of nodes
         color.border = "#013848", # border colour of nodes
         color.background = id.type, # background colour of nodes
         color.highlight.background = "#FF8000" # background colour of nodes when highlighted
         )

# plot graph
visNetwork(top1_betw_nodes, top1_betw_edges,
           height = "500px", width = "100%",
           main = paste0("Network Graph of ", top1_betw[["id"]])) %>%
  visIgraphLayout() %>%
  visGroups(groupname = "Beneficial Owner", shape = "triangle") %>%
  visGroups(groupname = "Beneficial Owner, Company Contacts", shape = "triangle") %>%
  visOptions(selectedBy = "group",
             highlightNearest = list(enabled = T, degree = 1, hover = T),
             nodesIdSelection = TRUE) %>% 
  visLayout(randomSeed = 123)
```

::: callout-note
## Key Observations

**Wave Warriors S.A. de C.V.** is a company node owned by many beneficial owners. There are also some cases where there are multiple layers of ownership. For example, we can see that **Matthew Williams** is a beneficial owner of **Wave Warriors S.A. de C.V.** and **Cole-Crosby** in turn is a beneficial owner of **Matthew Williams** (or his shareholdings).

Such behaviour of ownership layering is an anomaly as it could potentially be an attempt to obscure the ownership structure and make it challenging to trace the true ownership.
:::

## Analysis of Largest Connected Components

Next, we will perform an analysis on some of the largest connected components in the network chart. These large connected components can potentially reveal big groups of connected companies and beneficial owners. By identifying the nodes with the highest betweenness centrality in these components, we hope to be able to identify beneficial owners that are connected to many companies, which may be an anomaly or cause of concern. This is because when a beneficial owner controls multiple companies, it adds layers of complexity and obscurity to the ownership structure. Owning multiple companies can also provide a means to conceal illegal activities by interconnecting the companies or transferring funds or assets among the companies.

We will break up the network graph into connected components using the function [`components()`](https://search.r-project.org/CRAN/refmans/igraph/html/components.html) and find the three clusters that are largest in size.

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 14
#| fig-height: 14
# form graph
mc3_graph <- tbl_graph(nodes = mc3_nodes_master_updated,
                       edges = mc3_edges,
                       directed = FALSE)

# find components in graph
set.seed(123)
clusters <- components(mc3_graph)

# update graph with component membership
mc3_nodes_master_updated <- mc3_nodes_master_updated %>% 
  mutate(component_membership = clusters$membership)

# extract info relating to components
component_df <- clusters$csize %>% 
  as_tibble() %>% 
  rownames_to_column() %>% 
  rename(component_membership = rowname,
         component_size = value)

# find components that are top 3 in size    
top_3_components <- component_df %>% 
  top_n(3, component_size) %>% 
  arrange(desc(component_size))

datatable(top_3_components)
```

Next, we will visualise the network charts of the three largest clusters separately using interactive charts below.

::: callout-tip
## Interactive Chart

-   The darker orange nodes are the nodes with betweenness centrality above the 90th percentile in the network chart.
-   Select the id from drop-down or click on the interested entity on the network chart to view its first-degree linked nodes
-   Select the group from drop-down to view specific types of nodes. The group refers to the types of roles that the node may play in the entire network chart, e.g., Beneficial Owner or others
-   All nodes which play a role of Beneficial Owner in at least one of its relationships will be marked with a triangle shape. This makes it easier to identify multiple layers of beneficial owners where present.
:::

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 14
#| fig-height: 14
visualise_cluster <- function(x){
  
# extract nodes in component
component_nodes <- mc3_nodes_master_updated %>%
  filter(component_membership == x)

# extract edges in component
component_edges <- mc3_edges %>% 
  filter(source %in% component_nodes[["id"]] | target %in% component_nodes[["id"]])

# compute centrality measures
component_graph <- tbl_graph(nodes = component_nodes,
                             edges = component_edges,
                             directed = FALSE) %>% 
  mutate(closeness_centrality = centrality_closeness(),
         betweenness_centrality = centrality_betweenness(),
         eigen_cetrality = centrality_eigen())

# compute the top 90th percentile centrality
component_nodes_updated <- component_graph %>% 
  activate(nodes) %>% 
  as_tibble()

cent_per_90 <- quantile(component_nodes_updated$betweenness_centrality,
                               probs = 0.90)

component_nodes_updated <- component_nodes_updated %>% 
  mutate(is_top_cent_90 = ifelse(betweenness_centrality >= cent_per_90, "yes", "no"))

# colur palatte for betweenness centrality colours
sw_colors <- colorRampPalette(brewer.pal(3, "RdBu"))(3)

# customise edges for plotting
component_edges <- component_edges %>% 
  rename(from = source,
         to = target) %>% 
  mutate(title = paste0("Type: ", type), # tooltip when hover over
         color = "#0085AF") # color of edge

# customise nodes for plotting
component_nodes_updated <- component_nodes_updated %>% 
  rename(group = types) %>% 
  mutate(is_top_cent_90.type = ifelse(is_top_cent_90 == "yes", sw_colors[1], sw_colors[2])) %>% 
  mutate(title = paste0(id, "<br>Group: ", group), # tooltip when hover over
         size = 40, # set size of nodes
         color.border = "#013848", # border colour of nodes
         color.background = is_top_cent_90.type, # background colour of nodes
         color.highlight.background = "#FF8000" # background colour of nodes when highlighted
         )

# plot graph
visNetwork(component_nodes_updated, component_edges,
           height = "500px", width = "100%",
           main = paste0("Entities in Component ", x)) %>%
  visIgraphLayout() %>%
  visGroups(groupname = "Beneficial Owner", shape = "triangle") %>%
  visGroups(groupname = "Beneficial Owner, Company Contacts", shape = "triangle") %>%
  visGroups(groupname = "Beneficial Owner, Company", shape = "triangle") %>%
  visGroups(groupname = "Beneficial Owner, Company, Company Contacts", shape = "triangle") %>%
  visOptions(selectedBy = "group",
             highlightNearest = list(enabled = T, degree = 1, hover = T),
             nodesIdSelection = TRUE) %>% 
  visLayout(randomSeed = 123)

}

visualise_cluster(1)
```

::: callout-note
## Key Observations

This is a very large cluster with 7,093 nodes. From the network chart, we can see that the nodes are quite dispersed and there is no high density in any particular areas. As such, it is more difficult to single out certain nodes for observation. For further exploration, we can start with the darker orange nodes with betweenness centrality above the 90th percentile and find out anomalies from there.
:::

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 14
#| fig-height: 14
visualise_cluster(204)
```

::: callout-note
## Key Observations

This is the second largest cluster with 140 nodes. From this chart, we can identify some of the entities with high betweenness centrality as beneficial owners of many companies. For example, from the chart, we observe that **Mary Miller** is the beneficial owner of three entities, including **Andhra Pradesh Sp Worldwide**, **Henties Bay Ltd. Corporation**, **Mar del Este SE. Patterson**, **Fowler and Nicholson** on the other hand is linked to **Mary Miller** as Beneficial Owner, which could indicate that **Patterson**, **Fowler and Nicholson** is the ultimate beneficial owner of the three entities mentioned.

Such behaviour of ownership layering is an anomaly as it could potentially be an attempt to obscure the ownership structure and make it challenging to trace the true ownership.
:::

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 14
#| fig-height: 14
visualise_cluster(561)
```

::: callout-note
## Key Observations

This is the third largest cluster with 136 nodes. Similarly from this chart, we can identify some of the entities with high betweenness centrality as beneficial owners of many companies. For example, from the chart, we observe that **Jennifer Hill** is very well-connected, being the beneficial owner of two entities, including **Brisa del Mar GmbH** and **Mar de Cristal S.A. de C.V.** and the company contact of two entities, including **Coral Azul S.A. de C.V.** and **Branded tilapia Corporation Holdings**. This is potentially an anomaly as she holds control and influence over multiple companies and could potentially make use of the companies to engage in and conceal illegal activities.
:::

## Analysis of companies registered to many countries

Another anomaly is that we can look out for companies that are tied to or registered in many countries. This is an anomaly and risk because it canbe indicative of an attempt to take advantage of the regulatory systems in different jurisdictions to exploit loopholes or evade regulatory scrutiny. By frequently changing jurisdictions, the company may also make it more difficult for authorities to track its operations, identify connections to other entities, or scrutinise its financial transactions. It raises suspicions of trying to hide illicit activities or create a complex web of corporate structure.

We will view the top 3 companies that are linked to the most countries below. Specifically, the top company, **Aqua Aura SE Marine life** is found to be linked to nine countries, which is quite an anomaly. Coincidentally, **Aqua Aura SE Marine life** also falls under Cluster 1 that was the biggest connected component visualised above.

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 14
#| fig-height: 14
# extract node with most number of countries
top3_countries <- mc3_nodes_master_updated %>% 
  arrange(desc(num_countries)) %>% 
  top_n(3, num_countries) %>% 
  select(id, countries, types)

datatable(top3_countries)
```

We will further extract the first-degree and second-degree links of **Aqua Aura SE Marine life** for easier visualisation below.

::: callout-tip
## Interactive Chart

-   The darker orange node is our main node of interest, i.e., **Aqua Aura SE Marine life**
-   Select the id from drop-down or click on the interested entity on the network chart to view its first-degree linked nodes
-   Select the group from drop-down to view specific types of nodes. The group refers to the types of roles that the node may play in the entire network chart, e.g., Beneficial Owner or others
-   All nodes which play a role of Beneficial Owner in at least one of its relationships will be marked with a triangle shape. This makes it easier to identify multiple layers of beneficial owners where present.
:::

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 14
#| fig-height: 14
# extract node with most number of countries
top1_countries <- mc3_nodes_master_updated %>% 
  arrange(desc(num_countries)) %>% 
  top_n(1, num_countries) %>% 
  select(id, countries, types)

# extract lvl 1 edges
top1_countries_edges_lvl1 <- mc3_edges %>% 
  filter(source %in% top1_countries[["id"]] | target %in% top1_countries[["id"]])

# extract nodes from lvl 1 edges
id1 <- top1_countries_edges_lvl1 %>%
  select(source) %>%
  rename(id = source) %>% 
  left_join(mc3_nodes_master_updated, by = "id") %>% 
  select(id, countries, types)

id2 <- top1_countries_edges_lvl1 %>%
  select(target) %>%
  rename(id = target) %>% 
  left_join(mc3_nodes_master_updated, by = "id") %>% 
  select(id, countries, types)

additional_nodes_lvl1 <- rbind(id1, id2) %>% 
  distinct %>% 
  filter(!id %in% top1_countries[["id"]])

# extract lvl 2 edges
top1_countries_edges_lvl2 <- mc3_edges %>% 
  filter(source %in% additional_nodes_lvl1[["id"]] | target %in% additional_nodes_lvl1[["id"]])

# extract nodes from lvl 1 edges
id1 <- top1_countries_edges_lvl2 %>%
  select(source) %>%
  rename(id = source) %>% 
  left_join(mc3_nodes_master_updated, by = "id") %>% 
  select(id, countries, types)

id2 <- top1_countries_edges_lvl2 %>%
  select(target) %>%
  rename(id = target) %>% 
  left_join(mc3_nodes_master_updated, by = "id") %>% 
  select(id, countries, types)

additional_nodes_lvl2 <- rbind(id1, id2) %>% 
  distinct %>% 
  filter(!id %in% top1_countries[["id"]] & !id %in% additional_nodes_lvl1[["id"]])

# combine all nodes
top1_countries_nodes <- rbind(top1_countries, additional_nodes_lvl1, additional_nodes_lvl2) %>%
  distinct()

# combine all edges
top1_countries_edges <- rbind(top1_countries_edges_lvl1, top1_countries_edges_lvl2) %>% 
  distinct()

# colur palatte for betweenness centrality colours
sw_colors <- colorRampPalette(brewer.pal(3, "RdBu"))(3)

# customise edges for plotting
top1_countries_edges <- top1_countries_edges %>% 
  rename(from = source,
         to = target) %>% 
  mutate(title = paste0("Type: ", type), # tooltip when hover over
         color = "#0085AF") # color of edge

# customise nodes for plotting
top1_countries_nodes <- top1_countries_nodes %>% 
  rename(group = types) %>% 
  mutate(id.type = ifelse(id == top1_countries[["id"]], sw_colors[1], sw_colors[2])) %>% 
  mutate(title = paste0(id, "<br>Group: ", group), # tooltip when hover over
         size = 30, # set size of nodes
         color.border = "#013848", # border colour of nodes
         color.background = id.type, # background colour of nodes
         color.highlight.background = "#FF8000" # background colour of nodes when highlighted
         )

# plot graph
visNetwork(top1_countries_nodes, top1_countries_edges,
           height = "500px", width = "100%",
           main = paste0("Network Graph of ", top1_countries[["id"]])) %>%
  visIgraphLayout() %>%
  visGroups(groupname = "Beneficial Owner", shape = "triangle") %>%
  visGroups(groupname = "Beneficial Owner, Company Contacts", shape = "triangle") %>%
  visOptions(selectedBy = "group",
             highlightNearest = list(enabled = T, degree = 1, hover = T),
             nodesIdSelection = TRUE) %>% 
  visLayout(randomSeed = 123)
```

::: callout-note
## Key Observations

**Aqua Aura SE Marine life** is a company node owned by many beneficial owners. There are some cases where there are multiple layers of ownership. For example, we can see that **Michael Bell** is a beneficial owner of **Aqua Aura SE Marine life** and **Dean-Fisher** in turn is a beneficial owner of **Michael Bell** (or his shareholdings).

Such behaviour of ownership layering is an anomaly as it could potentially be an attempt to obscure the ownership structure and make it challenging to trace the true ownership.
:::

# References

-   [Modelling, Visualising and Analysing Network Data with R](https://r4va.netlify.app/chap27.html)
-   [Network Visualisation in R](https://jtr13.github.io/cc21fall2/network-visualization-in-r.html)
